---
title: "Assignment 3 Accidents"
author: "sai mithra"
date: "2023-10-15"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**QUESTION1**

##Our goal here is to predict whether an accident just reported will involve an injury (MAX_SEV_IR = 1 or 2) or will not (MAX_SEV_IR = 0). 
##For this purpose, create a dummy variable called INJURY that takes the value “yes” if MAX_SEV_IR = 1 or 2, and otherwise “no.”

```{r}
#load the complete data set and install/load required packages.

library(readr)
library(dplyr)

```

```{r}

library(caret)
library(e1071)

```

```{r}
Accidents_Data<- read.csv("C:\\Users\\saimi\\Downloads\\accidentsFull.csv")

View(Accidents_Data)

```

```{r}

#create and insert a dummy variable called injury in the data.

Accidents_Data$INJURY <- ifelse(Accidents_Data$MAX_SEV_IR>0, "yes", "no")

for (i in 1:dim(Accidents_Data)[2]) {
 if (is.character(Accidents_Data[, i])) {
 Accidents_Data[, i] <- as.factor(Accidents_Data[, i])
 }
}
head(Accidents_Data, n=24)

```

**QUESTION-1**

#Using the information in this dataset, if an accident has just been reported and no further information is available, what should the prediction be? (INJURY = Yes or No?) Why?

```{r}
#CREATING A TABLE BASED ON INJURY.

Injury_Table <- table(Accidents_Data$INJURY)
show(Injury_Table)

```
```{r}
#CALUCATING THE PROBABILITY OF INJURY 

Injury_Probablilty = 
scales::percent(Injury_Table["yes"]/(Injury_Table["yes"]+Injury_Table["no"]),
0.01)
Injury_Probablilty

```
**QUESTION-2**

#Select the first 24 records in the dataset and look only at the response (INJURY) and the two predictors WEATHER_R and TRAF_CON_R. 

```{r}
#make a new subset with just the necessary records in it.

Accidents_Data24 <- Accidents_Data[1:24, c('INJURY','WEATHER_R','TRAF_CON_R')] 
Accidents_Data24

```

#Create a pivot table that examines INJURY as a function of the two predictors for these 24 records. Use all three variables in the pivot table as rows/columns.

```{r}
dtpvt1 <- ftable(Accidents_Data24)
dtpvt2 <- ftable(Accidents_Data24 [,-1])

dtpvt1
dtpvt2

```
**Question-2(1)**

#Compute the exact Bayes conditional probabilities of an injury (INJURY = Yes) given the six possible combinations of the predictors.

```{r}

#QUESTION4
#COMPUTING THE BAYES CONDITIONAL PROBABLITIES OF AN INJURY (INJURY = Yes) GIVEN THE SIX POSSIBILE COMBINATIONS OF THE PREDITCTORS.

# Injury = yes

Prob1 = dtpvt1[3,1] / dtpvt2[1,1] # Injury, Weather=1 and Traf=0
Prob2 = dtpvt1[4,1] / dtpvt2[2,1] # Injury, Weather=2, Traf=0
Prob3 = dtpvt1[3,2] / dtpvt2[1,2] # Injury, W=1, T=1
Prob4 = dtpvt1[4,2] / dtpvt2[2,2] # I, W=2,T=1
Prob5 = dtpvt1[3,3] / dtpvt2[1,3] # I, W=1,T=2
Prob6 = dtpvt1[4,3]/ dtpvt2[2,3] #I,W=2,T=2
print(c(Prob1,Prob2,Prob3,Prob4,Prob5,Prob6))

# Injury = no

N1 = dtpvt1[1,1] / dtpvt2[1,1] # Weather=1 and Traf=0
N2 = dtpvt1[2,1] / dtpvt2[2,1] # Weather=2, Traf=0
N3 = dtpvt1[1,2] / dtpvt2[1,2] # W=1, T=1
N4 = dtpvt1[2,2] / dtpvt2[2,2] # W=2,T=1
N5 = dtpvt1[1,3] / dtpvt2[1,3] # W=1,T=2
N6 = dtpvt1[2,3] / dtpvt2[2,3] # W=2,T=2
print(c(N1,N2,N3,N4,N5,N6))


```

**QUESTION-2(2)**

#CLASSIFYING THE 24 ACCIDENTS USING THESES PROBABLITIES AND CUTOFF OF 0.5
#ADDING PROBABILITY RESULTS TO THE SUBSET 

```{r}
prob.inj <- rep(0,24)
for (i in 1:24) {
 print(c(Accidents_Data24$WEATHER_R[i],Accidents_Data24$TRAF_CON_R[i]))
 if (Accidents_Data24$WEATHER_R[i] == "1") {
 if (Accidents_Data24$TRAF_CON_R[i]=="0"){
 prob.inj[i] = Prob1
 }
 else if (Accidents_Data24$TRAF_CON_R[i]=="1") {
 prob.inj[i] = Prob3
 }
 else if (Accidents_Data24$TRAF_CON_R[i]=="2") {
 prob.inj[i] = Prob5
 }
 }
 else {
 if (Accidents_Data24$TRAF_CON_R[i]=="0"){
 prob.inj[i] = Prob2
 }
 else if (Accidents_Data24$TRAF_CON_R[i]=="1") {
 prob.inj[i] = Prob4
 }
 else if (Accidents_Data24$TRAF_CON_R[i]=="2") {
 prob.inj[i] = Prob6
 }
 }
}

```
```{r}
Accidents_Data24$prob.inj <- prob.inj
Accidents_Data24$pred.prob <- ifelse(Accidents_Data24$prob.inj>0.5, "yes", "no")
table(Accidents_Data24$pred.prob)
```

**QUESTION-2(3)**

#COMPUTING MANUALLY THE NAIVE BAYES CONDITIONAL PROBABILITY OF AN INJURY GIVEN THE WEATHER_R =1 AND TRAF_CON_R =1.

#The Naive Bayes conditional probability is computed using the Naive Bayes formula as follows:
#P(INJURY = Yes | WEATHER_R = 1 and TRAF_CON_R = 1) = (P(INJURY = Yes | WEATHER_R = 1) * P(INJURY = Yes | TRAF_CON_R = 1) * P(INJURY = Yes)) / (P(WEATHER_R = 1) * P(TRAF_CON_R = 1))

```{r}

Manual_NB_W1_T1 <- Prob3
cat("Manual Naive Bayes Conditional Probability (Injury = Yes | Weather_R = 
1, TRAF_CON_R = 1):", Manual_NB_W1_T1)

```
**QUESTION-3(4)**

#RUNNING A NAIVE BAYES CLASSIFIER ON THE 24 RECORDS AND TWO PREDICTORS.
#NOW,WE HAVE TO CHECK THE MODEL OUTPUT TO OBTAIN PROBABILITIES AND CLASSIFCATIONS FOR ALL 24 RECORDS.
##AND THEN, WE ARE COMPARING TO BAYES CLASSIFCATION TO SEE IF THE RESULTING CLASSIFICATIONS ARE EQUIVALENT OR NOT.

```{r}
library(e1071)

NB<-naiveBayes(INJURY ~ ., data = Accidents_Data24)

NBT <- predict(NB, newdata = Accidents_Data24,type = "raw")

Accidents_Data24$nbpred.prob <- NBT[,2] # Transfer the "Yes" nb prediction
library(caret)

NB2 <- train(INJURY ~ TRAF_CON_R + WEATHER_R, 
 data = Accidents_Data24, method = "nb")

predict(NB2, newdata = Accidents_Data24[,c("INJURY", "WEATHER_R", "TRAF_CON_R")])

predict(NB2, newdata = Accidents_Data24[,c("INJURY", "WEATHER_R", "TRAF_CON_R")],
 type = "raw")


```


**QUESTION-3**

#Let us now return to the entire dataset. Partition the data into training (60%) and validation (40%). 

```{r}

#Splitting the data into training (60%) and validation (40%)

set.seed(123)
TrainIndex <- createDataPartition(Accidents_Data$INJURY, p = 0.6, list =
FALSE)
Train_Data <- Accidents_Data[TrainIndex, ]
Val_Data <- Accidents_Data[-TrainIndex, ]

```

**QUESTION-3(1)**

Run a naive Bayes classifier on the complete training set with the relevant predictors (and INJURY as the response). Note that all predictors are categorical. Show the confusion matrix.

```{r}

#Creating a naive bayes model with the relavant predictors
nb <- naiveBayes(INJURY ~ WEATHER_R + TRAF_CON_R, data = Train_Data)
#Predicting on the validation set
Val_Pred <-predict(nb, newdata = Val_Data)

#Converting val_pred into a character vector
Val_Pred <- as.character(Val_Pred)

#Converting val_data$Injury to a character vector
Val_Data$INJURY <- as.character(Val_Data$INJURY)

#Creating a factor with matching levels
Val_Pred <- factor(Val_Pred, levels = c("No", "Yes"))
Val_Data$INJURY <- factor(Val_Data$INJURY, levels = c("No", "Yes"))

#Creating a confusion matrix
Confusion.Matrix <- confusionMatrix(Val_Pred, Val_Data$INJURY)
print(Confusion.Matrix)

```

**QUESTION-3(2)**

```{r}
#OVERALL ERROR OF THE VALIDATION SET 

Overall_Error <- 1 - Confusion.Matrix$overall["Accuracy"]
cat("overall error of the validation set:", Overall_Error, "\n")


```



